Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations

### 2024-09-08 Corrections_SAM_SS3_Krill_481.Rmd
add html

---
title: "Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "Appendix WP- SAM 2024/26"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The main recommendations were extracted from @CCAMLRWGSAM2024-

@@

@@

```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/compare4_Bratio_uncertainty.png')
```

```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/compare10_recruits_uncertainty.png')
```
## Observation 3.2

@@

Regarding stock unit issue, in the last online workshop of the SCAR Krill Expert Group (SKEG), conclude that The opinion of the SKEG was that the southern boundary of the ACC (sbACC) serves as an
important potential pathway for krill flux into Subarea 48.1, from the Bellingshausen Sea to the
northern Antarctic Peninsula, and into Subarea 48.2. The Weddell Front and Antarctic Slope
Front were also considered important pathways for krill flux into Subarea 48.2, but with focus in 48.1, with a importance of retention and flux (source to destination relation) identified by SKEG
should be taken into account when allocation of catch limits, particularly the importance
of retention in the Gerlach Strait, Bransfield Strait and Joinville Island strata, and the
directions of the important fluxes [WGEMM2439] (Figure \@ref(fig:ksh)).

@@

This consideration was implemented as a new scenario (`s9`) as previously described.

@@

```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
sspar(mfrow = c(5, 2), plot.cex = 0.8)
knitr::include_graphics(c('Figs/run1.png',
                          'Figs/run2.png',
                          'Figs/run3.png',
                          'Figs/run4.png',
                          'Figs/run5.png',
                          'Figs/run6.png',
                          'Figs/run7.png',
                          'Figs/run8.png',
                          'Figs/run9.png'))
```

To facilitate automated processing, results from several diagnostic tests can also be called as table. Table \@ref(tab:runtable)

```{r eval=TRUE, runtable, echo=FALSE}
data <- data.frame(
  Index = c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW", 
            "SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "PREDATOR"),
  runs.p = c(0.500, 0.145, 0.338, NA, 0.406, 0.189, 0.148, 0.334, 0.500, 0.599),
  test = c("Passed", "Passed", "Passed", "Excluded", "Passed", 
           "Passed", "Passed", "Passed", "Passed", "Passed"),
  sigma3.lo = c(-0.1816665, -0.2319052, -0.1813395, NA, -0.1476043, 
                -0.2452391, -0.2482065, -0.3723597, -0.5749614, -0.3154527),
  sigma3.hi = c(0.1816665, 0.2319052, 0.1813395, NA, 0.1476043, 
                0.2452391, 0.2482065, 0.3723597, 0.5749614, 0.3154527),
  type = rep("len", 10)
)

@@

Graph the Harvest Rate

```{r}
y <- base.model1$exploitation[,1]
x <- base.model1$exploitation[,4]

hr <- ggplot()+
  geom_line(aes(y,x))+
  geom_point(aes(y,x))+
  theme_bw()+
  labs(y="Harvest Rate",
       x ="")+
  ylim(0, 0.3)
hr
```

# SUPPLEMENTARY MATERIAL

- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)

- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)

- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)

>>>>

---
title: "Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "Appendix 1 WP- SAM 2024/26"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

@@

The main recommendations were extracted from @CCAMLRWGSAM2024.

@@

```{r include=FALSE, message=FALSE}
base.model1 <- SS_output(dir=dir1,
                        covar=T,
                        forecast=T)
```

@@

```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/COM1_9compare4_Bratio_uncertainty.png')
```

```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/COM1_9compare10_recruits_uncertainty.png')
```
## Observation 3.2

@@

Regarding stock unit issue, in the last online workshop of the SCAR Krill Expert Group (SKEG), conclude that The opinion of the SKEG was that the southern boundary of the ACC (sbACC) serves as an important potential pathway for krill flux into Subarea 48.1, from the Bellingshausen Sea to the northern Antarctic Peninsula, and into Subarea 48.2. The Weddell Front and Antarctic Slope Front were also considered important pathways for krill flux into Subarea 48.2, but with focus in 48.1, with a importance of retention and flux (source to destination relation) identified by SKEG should be taken into account when allocation of catch limits, particularly the importance of retention in the Gerlach Strait, Bransfield Strait and Joinville Island strata, and the directions of the important fluxes [@WGEMM2439] (Figure \@ref(fig:ksh)).

@@

This consideration was implemented as a new scenario (`s9`) as previously described and implemented in stock assessment of krill.

@@

```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
sspar(mfrow = c(5, 2), plot.cex = 0.8)
knitr::include_graphics(c('Figs/run1.png',
                          'Figs/run2.png',
                          'Figs/run3.png',
                          'Figs/run4.png',
                          'Figs/run5.png',
                          'Figs/run6.png',
                          'Figs/run7.png',
                          'Figs/run8.png',
                          'Figs/run9.png'))
```
To facilitate automated processing, results from several diagnostic tests can also be called as table. Table \@ref(tab:runtable)

```{r runtable, eval=TRUE, echo=FALSE}
data <- data.frame(
  Index = c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW", 
            "SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "PREDATOR"),
  runs.p = c(0.500, 0.145, 0.338, NA, 0.406, 0.189, 0.148, 0.334, 0.500, 0.599),
  test = c("Passed", "Passed", "Passed", "Excluded", "Passed", 
           "Passed", "Passed", "Passed", "Passed", "Passed"),
  sigma3.lo = c(-0.1816665, -0.2319052, -0.1813395, NA, -0.1476043, 
                -0.2452391, -0.2482065, -0.3723597, -0.5749614, -0.3154527),
  sigma3.hi = c(0.1816665, 0.2319052, 0.1813395, NA, 0.1476043, 
                0.2452391, 0.2482065, 0.3723597, 0.5749614, 0.3154527),
  type = rep("len", 10)
)

@@

Although it was not recorded in the WGSAM-2024 report, one of the questions raised was related to the representation of population variables, such as the harvest rate. In this document, we present them in Figure \@ref(fig:hr). for the entire analyzed series and for each stratum.

```{r hr, out.width='80%', fig.show='hold',  fig.cap="Harvest rate by strata in time series (1980-2020) in Antarctic krill fishey"}
# Convertir la matriz en un formato largo (long format) para ggplot
exploitation_long <- tidyr::pivot_longer(base.model1$exploitation, 
                                         cols = starts_with("FISHERY"), 
                                         names_to = "Fishery", 
                                         values_to = "Exploitation")

# Crear el gráfico
ggplot(exploitation_long, 
       aes(x = Yr, y = Exploitation, color = Fishery)) +
  geom_point() +
  geom_line() +
   geom_line(aes(y = annual_F), color = "black", linetype = "dashed", size = 1) + 
  scale_color_viridis_d()+
  labs(title = "",
       x = "Year",
       y = "Harvest Rat",
       color = "Strata") +
  theme_bw()+
  xlim(1980,2020)

```

### 2024-09-08 SS3_Krill_481_FSA.Rmd
add html

---
title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]

@@

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=TRUE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include those listed in Table \@ref(tab:datasources). These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2; 3). For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment. Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2). Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

@@

A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

@@

@@

```{r AKL, out.width='60%',  fig.cap="ALK Matrix Heatmap to krill in 48.1}
alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz

# Convertir la matriz en un data.frame
alk_df <- as.data.frame(as.table(alk_matrix))

# Renombrar columnas para mayor claridad
colnames(alk_df) <- c("Length", "TrueAge", "Value")

# Asegúrate de que las columnas Length y TrueAge sean numéricas
alk_df$Length <- as.numeric(as.character(alk_df$Length))
alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))

# Asegúrate de que la columna Value sea numérica
alk_df$Value <- as.numeric(alk_df$Value)

# Crear el plot usando ggplot
ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "True Age",
       y = "Length",
       fill = "Value") +
  theme_bw()+
  theme(legend.position = "none")
```

@@

@@

@@

For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).

@@

All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices. The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).

@@

@@

@@

The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021]. However, integrated models are not new to krill in SO. For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments. In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.

@@

@@

# FIGURES AND TABLES

```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.jpeg')
```

```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

```{r path, out.width='70%', fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
knitr::include_graphics('Figs/pathmod.png')
```

```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}

x_values <- seq(0, 7, by=0.1)
means <- c( 2.9,  3.5, 4.1)
std_devs <- c(0.65, 0.8, 0.65)
data <- data.frame(
  x = rep(x_values, each = length(means)),
  mean = rep(means, times = length(x_values)),
  std_dev = rep(std_devs, times = length(x_values))
)

```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
```

```{r res1a, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
```

\newpage
```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
```

```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
par(mfrow=c(2,2))
knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
                          'Figs/compare2_spawnbio_uncertainty.png',
                          'Figs/compare4_Bratio_uncertainty.png',
                          'Figs/compare17_densities_SSB_Virgin.png'))
```

>>>>

---
title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Antarctic Division, Southwest Fisheries Science Center, NOAA, USA]

@@

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=TRUE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      fig.show='hold',
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

@@

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include those listed in Table \@ref(tab:datasources). These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2; 3). For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.

Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2). Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

@@

A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population. There are two important issues related to the current stock assessment approach. The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult. However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean. Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1 [@WGEMM2439], however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the @CCAMLRWGSAM2024. Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

@@

\newpage 
```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}

### Age-Length Key Matrix

@@

```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
knitr::include_graphics('Figs/AKL.png')
```

@@

```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
x_values <- seq(0, 7, by=0.1)
means <- c( 2.9,  3.5, 4.1)
std_devs <- c(0.65, 0.8, 0.65)
data <- data.frame(
  x = rep(x_values, each = length(means)),
  mean = rep(means, times = length(x_values)),
  std_dev = rep(std_devs, times = length(x_values))
)

### Selectivity Blocks

According to the observations made at @CCAMLRWGSAM2024 regarding changes in net size and trawl type that could affect selectivity, we have incorporated a scenario (`s9`) with a selectivity block for krill. The inclusion of selectivity blocks aims to, on the one hand, reduce the effect of autocorrelation in the model fitting process, and on the other hand, explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms. According to the data identified, a shift from traditional to continuous trawling occurred in 2006. As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements. This is compared with the base scenario.

@@

```{r path, out.width='100%', fig.cap="Framework stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
knitr::include_graphics('Figs/pathmod.png')
```

It is important to note that there is a degree of consistency in how the performance of stock assessment models for toothfish and krill is evaluated. For krill, the primary diagnostics conducted adhere to the "best practices" outlined in @Carvalho2021b work. This recipe of steps for testing model performance is illustrated in Figure \@ref(fig:recipediags).

```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a base model."}
knitr::include_graphics('Figs/recipediags.png')
```

While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a converge towards these standard practices, regardless of the assessment platform and the structure of the integrated model.

@@

A comparative analysis of models `s01`, `s1`, and `s7` will be conducted, as they incorporate elements such as aggregate data, spatially structured data, and different sets of life history parameters according to @Mardones2024par. Additionally, to address the observations, scenarios `s8` and `s9` have been added and compared. The details of each observation made in the @CCAMLRWGSAM2024 report have been thoroughly addressed, and each response can be reviewed in **APPENDIX 1**.

@@

All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.

```{r res1,  out.width='60%', fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
```

```{r res1a, out.width='60%', fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
```

Regarding this pearson residual analysis, a visual inspection can lead to a mistaken conclusion about the model *goodness-of-fit*. We have integrated a quantitative element to evaluate this type of test. The plotting options are kept mainly to those provided by `r4ss`. With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model. Green shading indicates no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively. The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series. We plot this in Figure \@ref(fig:runtest) and Table \@ref(tab:runtable).

```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
sspar(mfrow = c(5, 2), plot.cex = 0.8)
knitr::include_graphics(c('Figs/run1.png',
                          'Figs/run2.png',
                          'Figs/run3.png',
                          'Figs/run4.png',
                          'Figs/run5.png',
                          'Figs/run6.png',
                          'Figs/run7.png',
                          'Figs/run8.png',
                          'Figs/run9.png'))
```

```{r runtable, eval=TRUE,  echo=FALSE}
data <- data.frame(
  Index = c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW", 
            "SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "PREDATOR"),
  runs.p = c(0.500, 0.145, 0.338, NA, 0.406, 0.189, 0.148, 0.334, 0.500, 0.599),
  test = c("Passed", "Passed", "Passed", "Excluded", "Passed", 
           "Passed", "Passed", "Passed", "Passed", "Passed"),
  sigma3.lo = c(-0.1816665, -0.2319052, -0.1813395, NA, -0.1476043, 
                -0.2452391, -0.2482065, -0.3723597, -0.5749614, -0.3154527),
  sigma3.hi = c(0.1816665, 0.2319052, 0.1813395, NA, 0.1476043, 
                0.2452391, 0.2482065, 0.3723597, 0.5749614, 0.3154527),
  type = rep("len", 10)
)

kable(data, format = "markdown", 
      col.names = c("Index", "runs.p", "test", "sigma3.lo", "sigma3.hi", "type"),
      caption = "Table 1: Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
  kable_styling(full_width = FALSE)
```

Regarding the fitting of indices in general, the model attempts to interpret the trends but encounters some difficulties in effectively fitting all sources of information, as shown in (Figure \@ref(fig:resall)).

```
The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:fitcom)).

@@

```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
```

@@

```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
par(mfrow=c(2,2))
knitr::include_graphics(c('Figs/COM1compare10_recruits_uncertainty.png',
                          'Figs/COM1compare2_spawnbio_uncertainty.png',
                          'Figs/COM1compare4_Bratio_uncertainty.png',
                          'Figs/COM1compare17_densities_SSB_Virgin.png'))
```

A selectivity block in an alternative scenario (`s9`) was implemented and compared its performance with the base model (`s1`). The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of recruit, spawning biomass, relative biomass and virgin biomass  (Figure \@ref(fig:comparisionselec)). It can be observed that there are no significant differences, at least in the state variables.

```{r comparisionselec, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
par(mfrow=c(2,2))
knitr::include_graphics(c('Figs/COM1_9compare10_recruits_uncertainty.png',
                          'Figs/COM1_9compare2_spawnbio_uncertainty.png',
                          'Figs/COM1_9compare4_Bratio_uncertainty.png',
                          'Figs/COM1_9compare17_densities_SSB_Virgin.png'))
```

@@

The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators. The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021]. However, integrated models are not new to krill in SO. For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments. In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*. Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is showed in Table \@ref(tab:parainit). However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed. This calculation process is specifically designed to avoid dependence on the initial conditioning parameters of the model.

@@

- This work is considered a continuous inductive analytical exercise, which takes into account elements influencing krill dynamics, such as data, factors, opinions, and others, and integrates them through a mathematical process. This exercise is iteratively refined with observations and corrections as it is communicated to the Antarctic community. Therefore, we appreciate the feedback provided in the WG-SAM and WG-EMM 2024, which has contributed to its improvement.

@@
